Geometric models reveal behavioral and neural signatures of transforming naturalistic experiences into episodic memories

### 2020-10-09 paper/main.tex
added responses to rev2 reviewers

Our work casts remembering as reproducing (behaviorally and neurally) the topic trajectory, or ``shape,'' of an experience.  When we characterized memory for a television episode using this framework, we found that every participant's recounting of the episode recapitulated the low spatial frequency details of the shape of its trajectory through topic space (Fig.~\ref{fig:trajectory}).  We termed this narrative scaffolding the episode's \textit{essence}.  Where participants' behaviors varied most was in their tendencies to recount specific low-level details from each episode event.  Geometrically, this appears as high spatial frequency distortions in participants' recall trajectories relative to the trajectory of the original episode (Fig.~\ref{fig:topics}).  We developed metrics to characterize the precision (recovery of any and all event-level information) and distinctiveness (recovery of event-specific information).  We also used word cloud visualizations to interpret the details of these event-level distortions.

>>>>

Our work casts remembering as reproducing (behaviorally and neurally) the topic trajectory, or ``shape,'' of an experience, thereby drawing implicit analogies between mentally navigating through word embedding spaces and physically navigating through spatial environments~\citep[e.g.,][]{BellEtal18, BellEtal20, ConsEtal16}.  When we characterized memory for a television episode using this framework, we found that every participant's recounting of the episode recapitulated the low spatial frequency details of the shape of its trajectory through topic space (Fig.~\ref{fig:trajectory}).  We termed this narrative scaffolding the episode's \textit{essence}.  Where participants' behaviors varied most was in their tendencies to recount specific low-level details from each episode event.  Geometrically, this appears as high spatial frequency distortions in participants' recall trajectories relative to the trajectory of the original episode (Fig.~\ref{fig:topics}).  We developed metrics to characterize the precision (recovery of any and all event-level information) and distinctiveness (recovery of event-specific information).  We also used word cloud visualizations to interpret the details of these event-level distortions.
